Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-28 Manuscript/manuscript.Rmd
fixed figure numbers

After obtaining our optimally built random forest classifier, we investigated whether our predictive models can alleviate resolution limitations of Hi-C data by predicting TAD boundaries at base pair resolution. We were able to identify regions where the model predicted the base pair to be a TAD boundary with probability equal to 1 (Figure 5). [??? Where is this figure? It seemed important; !!!added] By clustering the regions (see Methods), we were able to predict a comparable number of TAD boundary regions as called by the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line, respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 9). [??? Need further drill down, as we discussed; !!!added distance to annotations figure in next paragraph]

For each predicted TAD boundary region, we were able to identify the summit locations of TAD boundary location by performing partitioning around medoids (see Methods). To compare the summit locations of predicted and called TAD boundaries, we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flanking region. We found that the signal strength for the CTCF, RAD21, SMC3, and ZNF143 [??? Make all in this order!; !!!fixed] TFBS peaked at our predicted TAD boudaries, while was poorly defined around the called TAD boundaries (Figure 6; Supplementary Figure 10).  Thus, our model is able to alleviate resolution limitations and precisely locate TAD boundaries that correspond to the highest signal strength of known TAD boundary-forming TFBSs. Furthermore, when compared to called TAD boundaries, the boundaries that we predicted with our model were found to be statistically significantly closer to the top ranked TFBS, suggesting stronger enrichment profiles near predicted TAD boundaries (Supplementary Figure 11).

>>>>

After obtaining our optimally built random forest classifier, we investigated whether our predictive models can alleviate resolution limitations of Hi-C data by predicting TAD boundaries at base pair resolution. We were able to identify regions where the model predicted the base pair to be a TAD boundary with probability equal to 1 (Figure 5A). [??? Where is this figure? It seemed important; !!!added] By clustering the regions (see Methods), we were able to predict a comparable number of TAD boundary regions as called by the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line, respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 9). [??? Need further drill down, as we discussed; !!!added distance to annotations figure in next paragraph]

For each predicted TAD boundary region, we were able to identify the summit locations of TAD boundary location by performing partitioning around medoids (see Methods). To compare the summit locations of predicted and called TAD boundaries, we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flanking region. We found that the signal strength for the CTCF, RAD21, SMC3, and ZNF143 [??? Make all in this order!; !!!fixed] TFBS peaked at our predicted TAD boudaries, while was poorly defined around the called TAD boundaries (Figure 5B; Supplementary Figure 10).  Thus, our model is able to alleviate resolution limitations and precisely locate TAD boundaries that correspond to the highest signal strength of known TAD boundary-forming TFBSs. Furthermore, when compared to called TAD boundaries, the boundaries that we predicted with our model were found to be statistically significantly closer to the top ranked TFBS, suggesting stronger enrichment profiles near predicted TAD boundaries (Supplementary Figure 11).
